\documentclass[11pt,twoside]{report}
\usepackage[a4paper,width=150mm,top=25mm,bottom=25mm,bindingoffset=6mm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black,
    breaklinks=true
}
\usepackage{xstring}
\usepackage{adjustbox}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bbm}
\usepackage{bm}
\usepackage{booktabs}
\usepackage[style=numeric-comp, sorting=none]{biblatex}
\usepackage{braket}
\usepackage{catchfile}
\usepackage[capitalize]{cleveref}
\usepackage{color}
\usepackage{datatool}
\usepackage{easy-todo}
\usepackage{float}
\usepackage{sourcecodepro}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{multirow}
\usepackage{mathtools}
\usepackage{notoccite}
\usepackage{siunitx}
\usepackage{tabularx}
\usepackage{pythonhighlight}


\addbibresource{../references.bib}
\graphicspath{ {../images/}{../../results/plots/} }
\lstset{basicstyle=\small\ttfamily}

\input{../newcommands.tex}

%----------------------------------------------------------------------------------------
% Title Page
%----------------------------------------------------------------------------------------
\pagenumbering{gobble}
\title{
    {Quantum Boltzmann Machines}\\
    {\large Applications in Quantitative Finance}
}
\author{
    {\LARGE Cameron Perot\vspace{1cm}}\\
    {Master's Thesis\vspace{0.1cm}}\\
    {\small submitted to\vspace{0.1cm}}\\
    {the Faculty of Mathematics, Computer Science, and Natural Sciences}\\
    {of RWTH Aachen University\vspace{0.1cm}}\\
    {\small written at\vspace{0.1cm}}\\
    {Jülich Supercomputing Centre}\\
    {Forschungszentrum Jülich\vspace{1cm}}\\
    {First Examiner: Prof. Dr. Kristel Michielsen\( ^{*,\dag} \)}\\
    {Second Examiner: Prof. Dr. Holger Rauhut\( ^* \)}\\
    {Adviser: Dr. Dennis Willsch\( ^\dag \)\vspace{0.1cm}}\\
    {\footnotesize\( ^* \)RWTH Aachen University, D-52056 Aachen, Germany}\\
    {\footnotesize\( ^\dag \)Jülich Supercomputing Centre, Institute for Advanced Simulation,}\\
    {\footnotesize Forschungszentrum Jülich, D-52425 Jülich, Germany\vspace{0.5cm}}
}
\date{Date TBD}


\begin{document}
\maketitle
\pagenumbering{roman}

%----------------------------------------------------------------------------------------
% Abstract
%----------------------------------------------------------------------------------------
\clearpage\shipout\null
\chapter*{Abstract}

In this thesis we explore using the D-Wave Advantage 4.1 quantum annealer to sample quantum Boltzmann distributions and train quantum Boltzmann machines (QBMs).
We focus on the real-world problem of using QBMs as generative models to produce synthetic foreign exchange market data and analyze how the results stack up against classical models based on restricted Boltzmann machines.
Additionally, we study a small 12-qubit problem which we use to compare samples obtained from the annealer to theory, and in the process gain vital insights into how well the Advantage 4.1 can sample quantum Boltzmann random variables and be used to train QBMs.
Through this we are able to show that the D-Wave Advantage 4.1 can sample classical Boltzmann random variables to some extent, but is limited in its ability to sample from quantum Boltzmann distributions.
Our findings indicate that models trained using the annealer are much noisier than simulations and struggle to perform at the same level as classical models.

\clearpage\shipout\null

%----------------------------------------------------------------------------------------
% Table of Contents
%----------------------------------------------------------------------------------------
\tableofcontents
\clearpage\shipout\null

%----------------------------------------------------------------------------------------
% Introduction
%----------------------------------------------------------------------------------------
\chapter{Introduction}
\label{ch:introduction}
\pagenumbering{arabic}
\input{chapters/introduction.tex}

%----------------------------------------------------------------------------------------
% Data Analysis
%----------------------------------------------------------------------------------------
\chapter{Data Analysis \& Preprocessing}
\label{ch:data_analysis}
\input{chapters/data_analysis.tex}

%----------------------------------------------------------------------------------------
% RBM
%----------------------------------------------------------------------------------------
\chapter{Restricted Boltzmann Machine}
\label{ch:rbm}
\input{chapters/rbm.tex}

%----------------------------------------------------------------------------------------
% QBM
%----------------------------------------------------------------------------------------
\chapter{Quantum Boltzmann Machine}
\label{ch:qbm}
\input{chapters/qbm.tex}

%----------------------------------------------------------------------------------------
% Conclusion
%----------------------------------------------------------------------------------------
\chapter{Conclusion}
\label{ch:conclusion}
\input{chapters/conclusion.tex}

%----------------------------------------------------------------------------------------
% Appendix
%----------------------------------------------------------------------------------------
\appendix
\chapter{Appendix}
\label{ch:appendix}
\input{chapters/appendix.tex}

%----------------------------------------------------------------------------------------
% References
%----------------------------------------------------------------------------------------
\printbibliography

\end{document}
